Solving GENOPT problems \newline with the use of ExaMin solver

### 2016-05-10 2016/LION10/GenOpt.tex
Add final version of LION10 paper

\title{Solving GENOPT problems \newline with the use of ExaMin solver}
\author{Konstantin Barkalov (\email{konstantin.barkalov@itmm.unn.ru}[\Letter]) \and Alexander Sysoyev (\email{sysoyev@vmk.unn.ru})  \and Ilya Lebedev \email{lebedev.ilya.g@gmail.com } \and Vladislav Sovrasov (\email{sovrasov.vlad@gmail.com}) }

@@

This approach to the comparison of the algorithms has been used by the organizers of GENeralization-based contest in global OPTimization (GENOPT). The functions to be optimized are broadly divided into three function families: GKLS, conditioned transforms of classical benchmarks (Rosenbrock (unimodal, narrowing bending valley), Rastrigin (strongly multimodal), Zakharov (unimodal)), and a composition of classical benchmarks. In their turn, each family is subdivided into six function types, which differ by their analytical definition and by the number of dimensions. The functions belonging to the same type share the majority of properties and can be assumed to have the same difficulty. Finally, every function type of every family is realized into an unlimited number of function instances which differ by some randomly generated parameters. In particular, every instance will have a randomly generated offset $c\in[-1,1]$ added to the function’s output in order to make the global minimum value unpredictable. A detailed description of the competition problems can be found at the website \url{http://genopt.org}.

@@

Let us illustrate the work of this \textit{global search algorithm} (GSA) for minimization of a multiextremal function of two variables generated by GKLS generator \cite{Gaviano}. %function No 4, simple
The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=10$. Fig.~\ref{fig:1} shows the level lines of the function and the points of 421 trials carried out by the method before the required accuracy was obtained.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
  \caption{Minimization of a GKLS function of two variables by GSA}
  \label{fig:1}       % Give a unique label
\end{figure}

@@

The organizers of GENOPT competition have offered 18 classes of problems ($n=0,1,2$, see table \ref{tab:problems}). In each class, 100 functions have been given.

@@

The limit of 1 million computations of the function (trials) imposed by the organizers is essential for finding the global minimum of the functions with given dimensionalities. Thus, even for the 10-dimensional problems, having built a uniform grid with four points in each dimension, one goes already out of the 1 million points limit. The efficient algorithms, building the grids not dense everywhere, place the trials much more efficiently. However, for the 30-dimensional problems such an efficiency appears to be not enough as well. In this section, we describe four modifications of the core method used in solving various problems within the competition. The first two modifications have been applied to all problems, the third one – to the problems of GKLS class, and the last one has been used for solving the problems with Rastrigin, Rosenbrock, and Zakharov functions.

@@

One scheme aiming to accelerate the search process is to be outlined in this subsection (other schemes described in \cite{Kvasov2006,Kvasov2015} ). The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end-points of the corresponding interval. Following this idea, we introduce the modified characteristics 
\begin{equation}\label{eq:localR}
  R_\alpha(i) = \frac{R(i)}{\sqrt{(z_i-\varphi_k^\ast)(z_{i-1}-\varphi_k^\ast)}+\mu (1.5)^{-\alpha}},
\end{equation}
where $\mu$, $R(i)$ and $\varphi_k^\ast$ are respectively from (\ref{eq:11}), (\ref{eq:14}), (\ref{eq:phistar}), and $\alpha$ is some integer parameter for setting the desired level of localization.

@@

The modification with the separable stage with the precision $0.02$ for the stop condition of one-dimensional GSA was used for all classes except GKLS. When solving the GKLS problems, the modifications from subsections \ref{mixed} and \ref{hooke} have been used first, then modification from subsection \ref{random} was applied. The distribution of the number of the solved tasks with the use of random search and without the one is presented in table \ref{tab:randsolved}.

\begin{table}
	\caption{Total number of solved tasks from different classes}
	\label{tab:randsolved}
	\center
	\begin{tabular}{ccc}
		\hline\noalign{\smallskip}
	Class &  \parbox[c]{3cm}{ \centering  Solved without\\random mod}   & \parbox[c]{3cm}{ \centering Solved with\\random mod}    \\
	\noalign{\smallskip} \hline \noalign{\smallskip}		
		GKLS--nd--10 & 78 & 99 \\
		GKLS--nd--30 & 0 & 15 \\
		GKLS--cd--10 & 67 & 96 \\
		GKLS--cd--30 & 0 & 1 \\
		GKLS--td--10 & 65 & 94 \\
		GKLS--td--30 & 0 & 0 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\section{Conclusion}
In this work, the results of solving of the 10- and 30-dimensional problems of the unconditional global optimization from GenOpt 2016 competition are presented. The optimization methods used and the modifications of these ones directed onto the obtaining of a solution at given limitation of 1 million trials (computations of the objective function) are described. All the modifications considered are implemented in ExaMin solver employed in the conducting of the experiments. The numerical experiments have been carried out using Lobachevsky supercomputer \cite{Lobach}.\\

@@

\textbf{Acknowledgements.}This study was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''.

>>>>

\title{Solving GENOPT problems \newline with the use of ExaMin solver}
\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and \\ Ilya Lebedev \and Vladislav Sovrasov \\
\email{ \{konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\\vladislav.sovrasov\}@itmm.unn.ru}}

%\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and \newline Ilya Lebedev \and Vladislav Sovrasov \email{[konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\newline vladislav.sovrasov]@itmm.unn.ru}}

%\author{Konstantin Barkalov \Letter \and Alexander Sysoyev \and Ilya Lebedev \and Vladislav Sovrasov \email{[konstantin.barkalov,alexander.sysoyev,ilya.lebedev,\newline vladislav.sovrasov]@itmm.unn.ru}}

%\author{Konstantin Barkalov ( \Letter~\email{konstantin.barkalov@itmm.unn.ru})  \and \\ Alexander Sysoyev (\email{alexander.sysoyev@itmm.unn.ru}) \and \\ Ilya Lebedev (\email{ilya.lebedev@itmm.unn.ru})  \and \\ Vladislav Sovrasov (\email{vladislav.sovrasov@itmm.unn.ru}) }

@@

Approach to the comparison of the algorithms by solving a set of test problems has been used by the organizers of GENeralization-based contest in global OPTimization (GENOPT). The functions to be optimized are broadly divided into three function families: GKLS, conditioned transforms of classical benchmarks (Rosenbrock (unimodal, narrowing bending valley), Rastrigin (strongly multimodal), Zakharov (unimodal)), and a composition of classical benchmarks. In their turn, each family is subdivided into six function types, which differ by their analytical definition and by the number of dimensions. The functions belonging to the same type share the majority of properties and can be assumed to have the same difficulty. Finally, every function type of every family is realized into an unlimited number of function instances which differ by some randomly generated parameters. In particular, every instance will have a randomly generated offset $c\in[-1,1]$ added to the function’s output in order to make the global minimum value unpredictable. A detailed description of the competition problems can be found at the website \url{http://genopt.org}.

@@

%new
This \textit{global search algorithm} (GSA) was developed in the framework of information - statistical approach (see \cite{Strongin2000}). From this point of view the normalized characteristics $R(i)$ from (\ref{eq:14}) can be considered as probabilities of locating the global minimum within the interval  $(x_{i-1},x_i), 1\leq i \leq k+1$. Thus, at each iteration a new trial point is selected inside the interval, which has the greatest probability of finding the global minimum.
At the same time, use of running estimates (\ref{eq:11}) of Lipschitz constant $L$ allows us to apply GSA to the problems with a priori unknown values of $L$.

Let us illustrate the work of GSA for minimization of a multiextremal function of two variables generated by GKLS generator \cite{Gaviano}. %function No 4, simple
The experiment used the method parameters $r=3$, $\epsilon=10^{-3}$, and the evolvent construction parameter $m=10$. Fig.~\ref{fig:1} shows the level lines of the function and the points of 421 trials carried out by the method before the required accuracy was obtained.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig2.jpg} 
  \caption{Minimization of a GKLS function of two variables by GSA}
  \label{fig:1}       % Give a unique label
\end{figure}

@@

The organizers of GENOPT competition have offered 18 classes of problems (see table \ref{tab:problems}, %new
where last two rows correspond to six classes of composite functions with $n=0,1,2$). In each class, 100 functions have been given.

@@

The limit of 1 million computations of the function (trials) imposed by the organizers is essential for finding the global minimum of the functions with given dimensionalities. Thus, even for the 10-dimensional problems, having built a uniform grid with four points in each dimension, one goes already out of the 1 million points limit. %new
The global optimization algorithms that build non-uniform grid in the search domain put the points more efficiently. %new
However, for the 30-dimensional problems such an efficiency appears to be not enough as well. In this section, we describe four modifications of the core method used in solving various problems within the competition. The first two modifications have been applied to all problems, the third one – to the problems of GKLS class, and the last one has been used for solving the problems with Rastrigin, Rosenbrock, and Zakharov functions.

@@

One scheme aiming to accelerate the search process is to be outlined in this subsection (other schemes are described, for example, in \cite{Kvasov2006,Kvasov2015} ). The idea of acceleration is to magnify the characteristics of the intervals containing the best current estimates by introducing some factor depending on the function values estimated at the end-points of the corresponding interval. Following this idea, we introduce the modified characteristics 
\begin{equation}\label{eq:localR}
  R_\alpha(i) = \frac{R(i)}{\sqrt{(z_i-\varphi_k^\ast)(z_{i-1}-\varphi_k^\ast)}+\mu (1.5)^{-\alpha}},
\end{equation}
where $\mu$, $R(i)$ and $\varphi_k^\ast$ are respectively from (\ref{eq:11}), (\ref{eq:14}), (\ref{eq:phistar}), and $\alpha$ is some integer parameter for setting the desired level of localization.

@@

The modification with the separable stage with the precision $0.02$ for the stop condition of one-dimensional GSA was used for all classes except GKLS. When solving the GKLS problems, the modifications from subsections \ref{mixed} and \ref{hooke} have been used first, then modification from subsection \ref{random} was applied. %new
The number of the solved problems with and without random search is presented in table \ref{tab:randsolved}.%new

\begin{table}
	\caption{Total number of solved tasks from different classes}
	\label{tab:randsolved}
	\center
	\begin{tabular}{ccc}
		\hline\noalign{\smallskip}
	Class &  \parbox[c]{3cm}{ \centering  Solved without\\random mode}   & \parbox[c]{3cm}{ \centering Solved with\\random mode}    \\
	\noalign{\smallskip} \hline \noalign{\smallskip}		
		GKLS--nd--10 & 78 & 99 \\
		GKLS--nd--30 & 0 & 15 \\
		GKLS--cd--10 & 67 & 96 \\
		GKLS--cd--30 & 0 & 1 \\
		GKLS--td--10 & 65 & 94 \\
		GKLS--td--30 & 0 & 0 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\section{Conclusion}
In this work, the results of solving of the 10- and 30-dimensional problems of the unconditional global optimization from GenOpt 2016 competition are presented. The optimization methods used and the modifications of these ones directed onto the obtaining of a solution at given limitation of 1 million trials (computations of the objective function) are described. All the modifications considered are implemented in ExaMin solver employed in the conducting of the experiments. The numerical experiments have been carried out using Lobachevsky supercomputer \cite{Lobach}.

@@

\textbf{Acknowledgements.} This study was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''.
